Improving content discovery through combining linked data and data mining techniques

### 2016-05-06 report/chapters/06-evaluation.tex
Spell check

A higher-level
goal of this research was to perform some more qualitative evaluation
of those clusters by obtaining human feedback on a potential
application thereof. The application evaluted was a feature that would
take a page a user is currently reading and suggest related content
based on their being members of the same cluster. Mock-ups of such
a feature were generated and evaluated via a survery as described
in section~\ref{sec:survery}.

@@

The survey was taken by 30 respondants with three having taken it
following the think-aloud protocol\cite{lewis1982using} so that more
information can be gathered around the motivation for the responses.
Two of the think-aloud participants were experts in usability testing
and user experience design and the remaining participant a normal
website user.

@@

A key metric generated from the responses was the proportion of times
respondants selected a ``related'' set of items based on clusters
generated from a particular data pipeline variant. A table of the
responses is given in table~\ref{tbl:results}.

@@

The edges of the graph represent introducing one new technique and
re-running the experiment. For example, the edge from \emph{RDFa} to
\emph{RDFa+} shows what happens to the performance if we introduce
enrichment via derefence to data extracted also by dereference.
The edges from e.g. \emph{Entities} and \emph{Links} point to
the result of having used both of those techniques in data extraction
(joining the graphs via set union).

@@

In the particular results depicited, there are a mix of insignificant
and significant changes in performance when we alter one aspect of
the system. Insignificant results include:

@@

\begin{itemize}
\item Extraction via dereference only improves substantially when
  improved in \emph{any} way. It benefits most significantly out of
  all variants from enrichment.
\item Entity extraction benefits somewhat being combined with
  embedded semantics obtained via dereference, certainly more than
  it benefits from being combined with hyperlink relationships.
\item Hyperlink relationships perhaps benefit a little from
  enrichment (but it might be bold to claim that), but more significantly
  we can see that extraction technique in performence when combined
  with either of the two other extraction approaches (or both).
\end{itemize}

@@

Much of the performance gain by altering extraction via derefence
might be explainable purely due to having added more data properties
with which the machine learning algorithms can work. This starts to
suggest a potential for a minimum feature set size below which
clustering processes will struggle.

@@

The basic categories seemed to be powerful enough to please many
respondants who frequently approved of related content suggestions on
the basis that the suggestions were from the same part as the BBC
website. This may be an artefact of framing the survey in an
artificial way as people validating the system by whether the pages
are grouped by well-understand caterogies (e.g. football news, comedy
TV, science articles) does not necessarily correlate with how those
same users would have been tempted to follow links to suggested items
as part of their own casual browsing of the website. This point
was also raised in some of the think-aloud discussions with user
experience experts, as discussed in section~\ref{sec:eval-obs}.

@@

It is also important to note conclusions from the lack of significant
step changes and the most interesting one in these results is the lack
of support for enrichment via dereference improving performance at
all. This follows on from analysis in section~\ref{sec:anal-enrichment}
that suggests the use of derefence simply falls down on the same
drawbacks as dereference as an extraction technique. The only case
where enrichment makes a large difference is improving extraction via
dereference, but this could be explained more by the latter performing
so poorly in the first place, as discussed above.

@@

Further qualitative feedback was obtained in two ways: firstly by
adding to the end of the survery some open-ended questions to invite
some discussion:

@@

Approximately half of respondants answered the first two and only
three out of the 30 went on to give some additional comments. Many
of those were user experience experts who felt they had much to
contribute additional feedback.

@@

In the think-aloud observations, non-expert responses were terser
than those from experts and it seemed as though conclusions were
reached more quickly about
whether suggested content was relevant or related. The impression was
that non-experts may act on some instinct that is hard to
describe, whereas experts in user design are more experienced in
some of the rationales and patterns they have observed across many
user testing sessions. This more instinct-driven approach for
non-experts is supported further from observations that thumbnail
images were cited on more than one occassion as a motivator for
determining content as being similar or not.

The expert think-aloud observations highlighted the key difference
between content being similar due to being in the same categorisation
groups versus being related by themes or topics, as discussed at the
end of the previous section. The opinions from these user experience
experts was that a topic-based similarity would be strongest for users
to consider clicking through to the suggested content and this was
vindicated on a number of occassions in non-expert responses two
(in one example a respondant was disappointed that many suggestions
offered next to a BBC News article were just ``more articles'' and
not ``articles on the same topic'').

@@

The expert discussions raised some cases where ``hard'' categories
for the content were necessary, however. In some cases the clustering
processes had grouped content suitable for children with content not
explicitly marked as such. For a media organisation like the BBC, where
there is wide provision of content targed at children, any
promotional feature on the website needs to be very cautious about
what it suggests. In the BBC, it is typical to create a strict
separation between child and non-child content and a general-purpose
data extraction pipeline has no automatic knowledge of this. In a
real setting, it is likely a production version of this system would
want to have some basic business rules to constrain its operation
where these strict requirements exist.

@@

The final case where strict separation of content is needed is where
the system chose to cluster many British football teams together.
Semantically, the system was behaving very well in identifty that
all pages are published by BBC Sport and talk about a football team,
its fixtures and its results. However, with proper domain expertise,
a system should know not to suggest arbitrary, alternative football
teams to a user who is on the page for their favourite team. It is
unlikely they wish to read about teams they do not follow.

@@

\begin{itemize}
\item It might be that hundreds of objects linked via predicates
  creates too much noise in general. If those objects have come mainly
  from entity extraction, should entity extraction be tuned to bring
  back only the ``main'' topics of a page?
\item Chapter~\ref{chp:analysis} discussed how many of the extraction
  techniques (particularly deference) produced a lot of irrelevant
  facts (e.g. about the page itself, not the content) all of which
  would be candidiates for deference of the objects are IRIs.
  Effectively filtering out such noise would reduce the work spent
  enriching irrelevant things further.
\item A simple architectural solution is to parallelise the HTTP
  calls as much as possible such that an item's latency is only as
  long as the slowest response time. There are obvious limits to this
  (we should avoid overloading the server even if it is scaled for
  millions of visitors per day), but any parallelism is clearly
  better than a hundred sequential HTTP requests. Parallelism includes
  distributed architecture over multiple servers as much as
  concurrent programming paradigms.
\item An organisation should probably ensure a system is physically
  near any servers serving HTTP requests to reduce the overall TCP
  packet latency. This optimisation will add up over time, but is
  a challenge for a large organisation such as the BBC with systems
  of varying ages distributed over multiple locations
  \footnote{Industry has
  long outgrown the notion
  that enterprises like the BBC have a single ``www'' server as
  in the early days of the World Wide Web. However, the www hostname
  still needs to resolve to a single set of IP addresses for something
  that fronts the rest of the network like a load balancer. Effective
  caching in this front layer and ensuring our data pipline lives close
  to that cache might be sufficient in this case. An example of a
  front cache might be something like the Varnish HTTP cache:
  https://www.varnish-cache.org/}
\end{itemize}

Even with all the optimisations possible, it is unescapable that
multiple HTTP requests to assemble an RDF graph is slower than a
single query to a data store's query engine (e.g. SQL) that returns
all the information needed in a single response. The benefits of using
semantics, HTTP and other standards -- such as decoupling, simple
enrichment, reuse of standard technologies -- need to be effectively
researched and promoted, otherwise enterprise organisations will
continue to reach for the seemingly obvious solution of building
software that queries one database to push to another.

@@

Whilst there is much gained
from being able to reason about the entire breadth of an organisation's
content using semantics, it does not preclude writing bespoke business
rules that strip out known, erroneous data or provides a framework
against which to evaluate the likelihood of a fact being useful or
true. At the very least, there could be some work to develop ontologies
and schemas that detect inconsistencies in the data, but in a way
that adds value to a working system rather than such modelling being a
barrier to building the first iteration of the system.

@@

Other data attributes such as the Twitter
Cards metadata (see section~\ref{sec:smallest-cluster}) can contain
very broad things that are declared on all pages that use this
vocabulary (e.g. \texttt{card=summary} is always used to indicate that
a given page is using these meta tags) might cause a greey algorithm
to focus too long on grouping together all pages that happen to choose
to implement this Twitter feature with less concern about the more
interesting properties.

>>>>

A higher-level
goal of this research was to perform some more qualitative evaluation
of those clusters by obtaining human feedback on a potential
application thereof. The application evaluated was a feature that would
take a page a user is currently reading and suggest related content
based on their being members of the same cluster. Mock-ups of such
a feature were generated and evaluated via a survey as described
in section~\ref{sec:survery}.

@@

The survey was taken by 30 respondents with three having taken it
following the think-aloud protocol\cite{lewis1982using} so that more
information can be gathered around the motivation for the responses.
Two of the think-aloud participants were experts in usability testing
and user experience design and the remaining participant a normal
website user.

@@

A key metric generated from the responses was the proportion of times
respondents selected a ``related'' set of items based on clusters
generated from a particular data pipeline variant. A table of the
responses is given in table~\ref{tbl:results}.

@@

The edges of the graph represent introducing one new technique and
re-running the experiment. For example, the edge from \emph{RDFa} to
\emph{RDFa+} shows what happens to the performance if we introduce
enrichment via dereference to data extracted also by dereference.
The edges from e.g. \emph{Entities} and \emph{Links} point to
the result of having used both of those techniques in data extraction
(joining the graphs via set union).

@@

In the particular results depicted, there are a mix of insignificant
and significant changes in performance when we alter one aspect of
the system. Insignificant results include:

@@

\begin{itemize}
\item Extraction via dereference only improves substantially when
  improved in \emph{any} way. It benefits most significantly out of
  all variants from enrichment.
\item Entity extraction benefits somewhat being combined with
  embedded semantics obtained via dereference, certainly more than
  it benefits from being combined with hyperlink relationships.
\item Hyperlink relationships perhaps benefit a little from
  enrichment (but it might be bold to claim that), but more significantly
  we can see that extraction technique in performance when combined
  with either of the two other extraction approaches (or both).
\end{itemize}

@@

Much of the performance gain by altering extraction via dereference
might be explainable purely due to having added more data properties
with which the machine learning algorithms can work. This starts to
suggest a potential for a minimum feature set size below which
clustering processes will struggle.

@@

The basic categories seemed to be powerful enough to please many
respondents who frequently approved of related content suggestions on
the basis that the suggestions were from the same part as the BBC
website. This may be an artefact of framing the survey in an
artificial way as people validating the system by whether the pages
are grouped by well-understand categories (e.g. football news, comedy
TV, science articles) does not necessarily correlate with how those
same users would have been tempted to follow links to suggested items
as part of their own casual browsing of the website. This point
was also raised in some of the think-aloud discussions with user
experience experts, as discussed in section~\ref{sec:eval-obs}.

@@

It is also important to note conclusions from the lack of significant
step changes and the most interesting one in these results is the lack
of support for enrichment via dereference improving performance at
all. This follows on from analysis in section~\ref{sec:anal-enrichment}
that suggests the use of dereference simply falls down on the same
drawbacks as dereference as an extraction technique. The only case
where enrichment makes a large difference is improving extraction via
dereference, but this could be explained more by the latter performing
so poorly in the first place, as discussed above.

@@

Further qualitative feedback was obtained in two ways: firstly by
adding to the end of the survey some open-ended questions to invite
some discussion:

@@

Approximately half of respondents answered the first two and only
three out of the 30 went on to give some additional comments. Many
of those were user experience experts who felt they had much to
contribute additional feedback.

@@

In the think-aloud observations, non-expert responses were terser
than those from experts and it seemed as though conclusions were
reached more quickly about
whether suggested content was relevant or related. The impression was
that non-experts may act on some instinct that is hard to
describe, whereas experts in user design are more experienced in
some of the rationales and patterns they have observed across many
user testing sessions. This more instinct-driven approach for
non-experts is supported further from observations that thumbnail
images were cited on more than one occasion as a motivator for
determining content as being similar or not.

The expert think-aloud observations highlighted the key difference
between content being similar due to being in the same categorisation
groups versus being related by themes or topics, as discussed at the
end of the previous section. The opinions from these user experience
experts was that a topic-based similarity would be strongest for users
to consider clicking through to the suggested content and this was
vindicated on a number of occasions in non-expert responses two
(in one example a respondent was disappointed that many suggestions
offered next to a BBC News article were just ``more articles'' and
not ``articles on the same topic'').

@@

The expert discussions raised some cases where ``hard'' categories
for the content were necessary, however. In some cases the clustering
processes had grouped content suitable for children with content not
explicitly marked as such. For a media organisation like the BBC, where
there is wide provision of content targeted at children, any
promotional feature on the website needs to be very cautious about
what it suggests. In the BBC, it is typical to create a strict
separation between child and non-child content and a general-purpose
data extraction pipeline has no automatic knowledge of this. In a
real setting, it is likely a production version of this system would
want to have some basic business rules to constrain its operation
where these strict requirements exist.

@@

The final case where strict separation of content is needed is where
the system chose to cluster many British football teams together.
Semantically, the system was behaving very well in identifying that
all pages are published by BBC Sport and talk about a football team,
its fixtures and its results. However, with proper domain expertise,
a system should know not to suggest arbitrary, alternative football
teams to a user who is on the page for their favourite team. It is
unlikely they wish to read about teams they do not follow.

@@

\begin{itemize}
\item It might be that hundreds of objects linked via predicates
  creates too much noise in general. If those objects have come mainly
  from entity extraction, should entity extraction be tuned to bring
  back only the ``main'' topics of a page?
\item Chapter~\ref{chp:analysis} discussed how many of the extraction
  techniques (particularly deference) produced a lot of irrelevant
  facts (e.g. about the page itself, not the content) all of which
  would be candidates for deference of the objects are IRIs.
  Effectively filtering out such noise would reduce the work spent
  enriching irrelevant things further.
\item A simple architectural solution is to parallelise the HTTP
  calls as much as possible such that an item's latency is only as
  long as the slowest response time. There are obvious limits to this
  (we should avoid overloading the server even if it is scaled for
  millions of visitors per day), but any parallelism is clearly
  better than a hundred sequential HTTP requests. Parallelism includes
  distributed architecture over multiple servers as much as
  concurrent programming paradigms.
\item An organisation should probably ensure a system is physically
  near any servers serving HTTP requests to reduce the overall TCP
  packet latency. This optimisation will add up over time, but is
  a challenge for a large organisation such as the BBC with systems
  of varying ages distributed over multiple locations
  \footnote{Industry has
  long outgrown the notion
  that enterprises like the BBC have a single ``www'' server as
  in the early days of the World Wide Web. However, the www hostname
  still needs to resolve to a single set of IP addresses for something
  that fronts the rest of the network like a load balancer. Effective
  caching in this front layer and ensuring our data pipeline lives close
  to that cache might be sufficient in this case. An example of a
  front cache might be something like the Varnish HTTP cache:
  https://www.varnish-cache.org/}
\end{itemize}

Even with all the optimisations possible, it is inescapable that
multiple HTTP requests to assemble an RDF graph is slower than a
single query to a data store's query engine (e.g. SQL) that returns
all the information needed in a single response. The benefits of using
semantics, HTTP and other standards -- such as decoupling, simple
enrichment, reuse of standard technologies -- need to be effectively
researched and promoted, otherwise enterprise organisations will
continue to reach for the seemingly obvious solution of building
software that queries one database to push to another.

@@

While there is much gained
from being able to reason about the entire breadth of an organisation's
content using semantics, it does not preclude writing bespoke business
rules that strip out known, erroneous data or provides a framework
against which to evaluate the likelihood of a fact being useful or
true. At the very least, there could be some work to develop ontologies
and schemas that detect inconsistencies in the data, but in a way
that adds value to a working system rather than such modelling being a
barrier to building the first iteration of the system.

@@

Other data attributes such as the Twitter
Cards metadata (see section~\ref{sec:smallest-cluster}) can contain
very broad things that are declared on all pages that use this
vocabulary (e.g. \texttt{card=summary} is always used to indicate that
a given page is using these meta tags) might cause a greedy algorithm
to focus too long on grouping together all pages that happen to choose
to implement this Twitter feature with less concern about the more
interesting properties.

### 2016-05-06 report/chapters/07-conclusions.tex
Spell check

\section{Challenges and Opportunitues for Semantics in the Enterprise}

As discussed in section~\ref{sec:production-recommendations} and also more
generally throughout chapter~\ref{chp:analysis}, there is much scope for
enterprises to benefit from using semantics in public-facing pages that display
their own content. This is particulary necessary in larger organisations where
attemping to gain an overview of all the content metadata in a short timeframe
would be a large enterprise integration effort.

@@

The development time on the extraction of embedded semantics is
far less than a series of bespoke enterprise integation efforts
against each data store and there are immediate results in some
of the clusters it can produce.

>>>>

\section{Challenges and Opportunities for Semantics in the Enterprise}

As discussed in section~\ref{sec:production-recommendations} and also more
generally throughout chapter~\ref{chp:analysis}, there is much scope for
enterprises to benefit from using semantics in public-facing pages that display
their own content. This is particularly necessary in larger organisations where
attempting to gain an overview of all the content metadata in a short timeframe
would be a large enterprise integration effort.

@@

The development time on the extraction of embedded semantics is
far less than a series of bespoke enterprise integration efforts
against each data store and there are immediate results in some
of the clusters it can produce.

### 2016-05-06 report/report.tex
Spell check

\title{Improving content discovery through combining linked data and data mining techniques}
\author{\vfill\textbf{Ross Fenning}}
\affil{
\vfill
  Dissertation submitted in partial fulfilment of the
\\
requirements for the degree of
\\
Master by Advanced Study in Software Engineering and Internet Architecture

@@

This project designs and implements a process by which semantic data
about any piece of online media content can be extracted into RDF
models in
multiple ways and feature
sets for machine learning can be generated from
that data. The clusters produced by fourteen different permutations
of techniques are evaluated and a potential application of suggesting
related content is mocked up for qualititative evaluation.

>>>>

\title{Improving content discovery through combining linked data and data mining techniques}
\author{\vfill\textbf{Ross Fenning}}
\affil{
\vfill
  Dissertation submitted in partial fulfillment of the
\\
requirements for the degree of
\\
Master by Advanced Study in Software Engineering and Internet Architecture

@@

This project designs and implements a process by which semantic data
about any piece of online media content can be extracted into RDF
models in
multiple ways and feature
sets for machine learning can be generated from
that data. The clusters produced by fourteen different permutations
of techniques are evaluated and a potential application of suggesting
related content is mocked up for qualitative evaluation.
